### 2018-02-03 submission/manuscript.tex
update to change CRC in specific places

\textbf{Background.} An increasing body of literature suggests that
there is a crucial role for the microbiota in colorectal cancer (CRC)
pathogenesis. Important drivers within this context have ranged from
individual microbes to the whole community. Our study expands on a
recent meta-analysis investigating microbial biomarkers for CRC by
testing the hypothesis that the bacterial community has important
associations to both early (adenoma) and late (carcinoma) stage disease.
To test this hypothesis we examined both feces (n = 1737) and colon
tissue (492 total samples from 350 individuals) across 14 previously
published 16S rRNA gene sequencing studies on CRC and the microbiota.

@@

Colorectal cancer (CRC) is a growing world-wide health problem in which
the microbiota has been purported to play an active role in disease
pathogenesis {[}1,2{]}. Numerous studies have shown the importance of
both individual microbes {[}3--7{]} and the overall community
{[}8--10{]} in tumorgenesis using mouse models of CRC. There have also
been numerous case/control studies investigating the microbiota in the
formation of both adenoma and carcinoma. A recent meta-analysis
investigated whether specific biomarkers could be consistently
identified using multiple data sets {[}11{]}. This meta-analysis focused
on identifying microbial signatures of CRC (biomarkers) but did so on a
small total number of individuals and only investigated stool. This
present meta-analysis addresses some of these major shortcomings.

Although there has been an intense focus on microbiota-based biomarker
discovery for CRC, the number of candidate genera seem to be endless.
Some studies point towards mouth-associated genera such as
\emph{Fusobacterium}, \emph{Peptostreptococcus}, \emph{Parvimonas}, and
\emph{Porphyromonas} as key enriched genera {[}6,12--18{]}. Yet, even in
these studies, mouth-associated genera are far from the only microbes
identified to be associated with CRC. These other genera include, but
are not limited to, \emph{Providencia}, \emph{Mogibacterium},
\emph{Enterococcus}, \emph{Escherichia/Shigella}, \emph{Klebsiella}, and
\emph{Streptococcus} {[}14--16{]}. In fact, there is good \emph{in vivo}
evidence that \emph{Escherichia/Shigella} and \emph{Streptococcus} can
be important in the pathogenesis of CRC {[}5,19,20{]}. Other studies
have also identified \emph{Akkermansia muciniphila} and
\emph{Bacteroides fragilis} as potential markers of CRC with good
mechanistic studies for the latter {[}15,21,22{]}. A recent
meta-analysis confirmed the correlations of certain mouth-associated
genera and \emph{Akkermansia muciniphila} with carcinoma {[}11{]}.
However, the sample size (n = 509) is equal to or less than some more
recent individual studies investigating the microbiota and CRC, making
it hard to know how extrapolatable these findings are. That particular
meta-analysis also added more potential microbial associations to both
carcinoma (\emph{Pantoea agglomerans} \emph{Ruminococcus},
\emph{Lactobacillus}) and adenoma (\emph{Prevotella},
\emph{Methanosphaera}, \emph{Succinovibrio}, \emph{Haemophilus
parainfluenzae}, \emph{Ruminococcus}, \emph{Lactobacillus}) stages of
CRC that need to be investigated further, since a number of these genera
have been found to be enriched in controls and not disease
{[}13,16,17{]}. Additionally, genera like \emph{Roseburia} have been
found in some studies to be increased in CRC but in others to either be
decreased or have no difference {[}15,18,23,24{]}.

Most of these studies have focused on carcinoma but the adenoma
observations are not any clearer at identifying candidate genera of
disease. Groups focusing on broad scale community metrics have found
that metrics such as richness are decreased in the adenoma stage of CRC
versus controls. Other studies have identified \emph{Lactococcus},
\emph{Pseudomonas}, \emph{Acidovorax}, \emph{Cloacibacterium},
\emph{Helicobacter}, \emph{Lactobacillus}, \emph{Bilophila},
\emph{Desulfovibrio}, and \emph{Mogibacterium} to be increased in
adenoma {[}25--27{]}. Based on the studies mentioned, there seems to be
very little overlap between the genera identified to be associated with
adenoma and carcinoma, with \emph{Lactobacillus} being one of the few
commonalities.

Targeting the identification of CRC microbial biomarkers within stool
seems logical since it offers an easy and cost-effective way to stratify
risk and the current gold standard for diagnosis, a colonoscopy, can be
time-consuming and is not without risk of complications. Although stool
represents an easy and less invasive way to assess risk, it is not clear
how well this sample reflects adenoma- and carcinoma- associated
microbial communities. Some studies have tried to assess this in health
and disease but are limited by their sample size {[}18,28{]}. Sampling
the microbiota directly associated with colon tissue may provide clearer
answers but is not without limitations. The community present for
sampling following the colonoscopy bowel prep may reflect the better
adhered microbiota versus the resident community. Additionally, these
samples contain more host DNA, potentially limiting the types of
analysis that can be done. It is well known that low biomass samples can
be very difficult to work with and results can be study dependent due to
the randomness of contamination {[}29{]}.

In comparison to the previous meta-analysis, this study significantly
increases the total stool samples investigated, re-examines important
genera across adenoma and carcinoma across study, and examines
differences and similarities between stool and tissue microbiota in the
context of CRC. Importantly, this analysis and approach could provide
valuable insights into the common genera that are both protective and
detrimental in CRC and whether broad bacterial community measurements
can account for these changes that were not provided by earlier
meta-analysis studies {[}11{]}.

Using both feces (n = 1737) and colon tissues (492 samples from 350
individuals) totaling over 2229 total samples across 14 studies
{[}12--18,21,23--27,30{]} {[}Table 1 \& 2{]}, we expand both the breadth
and scope of the previous meta-analysis to investigate whether the
bacterial community or specific members are more important risk factors
for both adenoma and carcinoma stages of CRC. To accomplish this we
first assessed whether bacterial diversity changes throughout disease
(control to adenoma to carcinoma) and if it results in an increased
relative risk (RR) for adenoma or carcinoma stages of CRC. We then
assessed what genera, if any, increase or decrease the RR of adenoma or
carcinoma stages of CRC. Next, using Random Forest models, we analyzed
whether the full community or only the combined top 5 increased and top
5 decreased RR genera resulted in better model classification, based on
the area under the curve (AUC). Finally, we also examined at what effect
and sample size the studies used were powered for and the sample size
needed to get to the traditionally accepted 80\% power. Our results from
these analyses suggests that the bacterial community changes as disease
severity worsens, that specific members are important for disease
classification, and that many of the studies are underpowered for
assessing small effect sizes.

@@

\textbf{\emph{Lower Bacterial Diversity is Associated with Increased RR
of CRC:}} To assess differences in broad scale community metrics as
disease severity worsens Operational Taxonomic Unit (OTU) richness,
evenness, and Shannon diversity measurements were power transformed and
Z-score normalized. These metrics are commonly used to assess the total
number of OTUs, the equality of their abundance, and the overall
diversity, respectively. Using linear mixed-effect models to control for
study and variable region we assessed whether OTU richness, evenness, or
Shannon diversity changed in a step-wise manner with disease severity.
In stool, there was a significant decrease in both evenness and Shannon
diversity as disease severity moved from control to adenoma to carcinoma
(P-value = 0.025 and 0.043, respectively) {[}Figure 1A{]}. We next
tested whether the detectable differences in community significantly
increased in RR of having an adenoma or carcinoma. For fecal samples, a
decrease versus the overall median in evenness resulted in a
significantly increased RR for carcinoma (RR = 1.36 (1.15 - 1.61),
P-value = 0.00034) and adenoma (RR = 1.16 (1.01 - 1.34), P-value =
0.032) while a decrease versus the overall median in Shannon diversity
only increased the RR for carcinoma (RR = 1.33 (1.09 - 1.62), P-value =
0.0047) {[}Figure 2{]}. Using the Bray-Curtis distance metric and
PERMANOVA, it was also possible to identify significant bacterial
community changes, in specific studies, for both carcinoma-associated
and adenoma-associated microbiota versus control {[}Table S1 \& S2{]}.

@@

\textbf{\emph{Mouth-Associated Genera are Associated with an Increased
RR of CRC:}} Next, we asked if being higher than the median relative
abundance, for any specific genera, resulted in an altered RR for
adenoma or carcinoma, in stool and colon tissue, due to our previous
observations of small increases in RR using OTU richness and Shannon
diversity. To investigate this we analyzed all common genera across each
study, in colon tissue or stool, and assessed whether a relative
abundance higher than the median results in an increase or decrease in
RR. Mouth-associated genera were commonly found in the top 5 genera
associated with an increased RR of having an adenoma
(\emph{Pyramidobacter} {[}Figure 3A{]} and \emph{Rothia} {[}Figure
3C{]}) and carcinoma (\emph{Fusobacterium}, \emph{Parvimonas},
\emph{Porphyromonas}, and \emph{Peptostreptococcus} {[}Figure 3B{]} and
\emph{Fusobacterium} {[}Figure 3D{]}) for both stool and colon tissue
samples. Conversely, genera commonly associated with a normal
gastrointestinal tract were correlated with a decreased RR for both
adenoma and carcinoma for both stool and colon tissue samples {[}Figure
3{]}. Even though mouth-associated genera were identified across disease
stage, there was little direct overlap of the top 5 increased or
decreased RR genera between both stages and sample site.

When observing RRs with a P-value less than 0.05 there was almost no
overlap between genera from stool or colon tissue and when they were
similar the RR was in opposite directions (e.g. \emph{Lactococcus})
{[}Table S8 \& S9{]}. Many of the genera that had RRs with a P-value
under 0.05 for colon tissue are also highly prevalent in contamination,
specifically, \emph{Novosphingobium}, \emph{Selemonas}, and
\emph{Achromobacter} {[}Figure 3 \& Table S8-S9{]}. For carcinoma stage
of CRC, certain mouth-associated genera (\emph{Fusobacterium},
\emph{Parvimonas}, \emph{Peptostreptococcus}) had a high RR for both
colon tissue and stool samples {[}Table S10 \& S11{]}. Finally, these
data suggest that the most significantly increased RR genera for tissue
was \emph{Camplyobacter} while in stool it was \emph{Peptostreptococcus}
{[}Table S10 \& S11{]}.

\textbf{\emph{Select Genera Models can Recapitulate Whole Community
Models:}} Since specific genera increased RR for carcinoma over
diversity metrics we assessed whether the bacterial community was better
at classifying disease versus only a select group of genera. We selected
these genera based on their RR and P-value significance and used two
approaches to test this question. The first approach used genus level
data and tested for differences in AUC between all genera and selected
genera. A single study was used for training the model prior to testing
on all other studies and this was repeated for every study in the
meta-analysis. The second approach used OTU level data and tested for a
generalized decrease in the 10-fold cross validation (CV) model AUC
which is a common approach used to guard against over-fitting. This was
applied across study and the AUC of the all OTUs model was compared
against the model that used only OTUs that taxonomically classified to
selected genera.

@@

Our study identifies clear differences in diversity, both at the
community level and for individual genera, present in patients with and
without CRC {[}Figure 1-3{]}. Although there was a step-wise decrease in
diversity as disease progressed from control to adenoma to carcinoma,
this did not translate into large effect sizes for the RR of CRC. Even
though mouth-associated genera increased the RR of having a carcinoma,
they did not consistently increase the RR of having an adenoma.
Additionally, our observations suggest that by combining
mouth-associated and CRC protective microbes we can classify either
adenoma or carcinoma stage of disease as well as models that use the
full community.

@@

The colon tissue-based studies did not provide a clearer understanding
of how the microbiota may be associated with CRC. Generally, the full
OTU-based models of unmatched and matched colon tissue samples were
concordant with stool samples showing that GI resident microbes were the
most prevalent in the top 10 most important variables across study
{[}Figure S7E \& F{]}. Unlike in stool, \emph{Fusobacterium} was the
only mouth-associated bacteria consistently present in the top 10 most
important variables of the full carcinoma stage models {[}Figure S7B-C
\& E-F{]}. The majority of the colon tissue-based results seem to be
study specific with many of the top 10 taxa being present only in a
single study. Additionally, the presence of genera associated with
contamination, within the top 10 most important variables for the genera
and OTU models is worrying. The low bacterial biomass of tissue samples
coupled with potential contamination could explain why these results
seem to be more sporadic than the stool results.

@@

\textbf{\emph{Matched versus Unmatched Tissue Samples:}} In general,
tissue samples with control and CRC samples from different individuals
were classified as unmatched while samples that belonged to the same
individual were classified as matched. Studies with matched data
included Burns, Dejea, Geng, and Lu while those with unmatched data were
from Burns, Flemer, Chen, and Sanapareddy. For some studies samples
became unmatched when a corresponding matched sample did not make it
through sequence processing. All samples, from both tissue sample types,
were analyzed together for the linear mixed-effect models with samples
from the same individual corrected for. For all other analysis, not
mentioned herein, matched and unmatched samples were analyzed separately
using the statistical approaches mentioned in the Statistical Analysis
section.

>>>>

\textbf{Background.} An increasing body of literature suggests that
there is a crucial role for the microbiota in colorectal cancer (CRC)
pathogenesis. Important drivers within this context have ranged from
individual microbes to the whole community. Our study expands on a
recent meta-analysis investigating microbial biomarkers for tumors by
testing the hypothesis that the bacterial community has important
associations to both early (adenoma) and late (carcinoma) stage disease.
To test this hypothesis we examined both feces (n = 1737) and colon
tissue (492 total samples from 350 individuals) across 14 previously
published 16S rRNA gene sequencing studies on CRC and the microbiota.

@@

Colorectal cancer (CRC) is a growing world-wide health problem in which
the microbiota has been purported to play an active role in disease
pathogenesis {[}1,2{]}. Numerous studies have shown the importance of
both individual microbes {[}3--7{]} and the overall community
{[}8--10{]} in tumorgenesis using mouse models of CRC. There have also
been numerous case/control studies investigating the microbiota in the
formation of both adenoma and carcinoma. A recent meta-analysis
investigated whether specific biomarkers could be consistently
identified using multiple data sets {[}11{]}. This meta-analysis focused
on identifying microbial signatures of tumors (biomarkers) but did so on
a small total number of individuals and only investigated stool. This
present meta-analysis addresses some of these major shortcomings.

Although there has been an intense focus on microbiota-based biomarker
discovery for tumors, the number of candidate genera seem to be endless.
Some studies point towards mouth-associated genera such as
\emph{Fusobacterium}, \emph{Peptostreptococcus}, \emph{Parvimonas}, and
\emph{Porphyromonas} as key enriched genera {[}6,12--18{]}. Yet, even in
these studies, mouth-associated genera are far from the only microbes
identified to be associated with CRC. These other genera include, but
are not limited to, \emph{Providencia}, \emph{Mogibacterium},
\emph{Enterococcus}, \emph{Escherichia/Shigella}, \emph{Klebsiella}, and
\emph{Streptococcus} {[}14--16{]}. In fact, there is good \emph{in vivo}
evidence that \emph{Escherichia/Shigella} and \emph{Streptococcus} can
be important in the pathogenesis of CRC {[}5,19,20{]}. Other studies
have also identified \emph{Akkermansia muciniphila} and
\emph{Bacteroides fragilis} as potential markers of CRC with good
mechanistic studies for the latter {[}15,21,22{]}. A recent
meta-analysis confirmed the correlations of certain mouth-associated
genera and \emph{Akkermansia muciniphila} with carcinoma {[}11{]}.
However, the sample size (n = 509) is equal to or less than some more
recent individual studies investigating the microbiota and colorectal
tumors, making it hard to know how extrapolatable these findings are.
That particular meta-analysis also added more potential microbial
associations to both carcinoma (\emph{Pantoea agglomerans}
\emph{Ruminococcus}, \emph{Lactobacillus}) and adenoma
(\emph{Prevotella}, \emph{Methanosphaera}, \emph{Succinovibrio},
\emph{Haemophilus parainfluenzae}, \emph{Ruminococcus},
\emph{Lactobacillus}) stages of disease that need to be investigated
further, since a number of these genera have been found to be enriched
in controls and not disease {[}13,16,17{]}. Additionally, genera like
\emph{Roseburia} have been found in some studies to be increased in
tumors but in others to either be decreased or have no difference
{[}15,18,23,24{]}.

Most of these studies have focused on carcinoma but the adenoma
observations are not any clearer at identifying candidate genera of
disease. Groups focusing on broad scale community metrics have found
that metrics such as richness are decreased in the adenoma stage of
disease versus controls. Other studies have identified
\emph{Lactococcus}, \emph{Pseudomonas}, \emph{Acidovorax},
\emph{Cloacibacterium}, \emph{Helicobacter}, \emph{Lactobacillus},
\emph{Bilophila}, \emph{Desulfovibrio}, and \emph{Mogibacterium} to be
increased in adenoma {[}25--27{]}. Based on the studies mentioned, there
seems to be very little overlap between the genera identified to be
associated with adenoma and carcinoma, with \emph{Lactobacillus} being
one of the few commonalities.

Targeting the identification of tumor microbial biomarkers within stool
seems logical since it offers an easy and cost-effective way to stratify
risk and the current gold standard for diagnosis, a colonoscopy, can be
time-consuming and is not without risk of complications. Although stool
represents an easy and less invasive way to assess risk, it is not clear
how well this sample reflects adenoma- and carcinoma- associated
microbial communities. Some studies have tried to assess this in health
and disease but are limited by their sample size {[}18,28{]}. Sampling
the microbiota directly associated with colon tissue may provide clearer
answers but is not without limitations. The community present for
sampling following the colonoscopy bowel prep may reflect the better
adhered microbiota versus the resident community. Additionally, these
samples contain more host DNA, potentially limiting the types of
analysis that can be done. It is well known that low biomass samples can
be very difficult to work with and results can be study dependent due to
the randomness of contamination {[}29{]}.

In comparison to the previous meta-analysis, this study significantly
increases the total stool samples investigated, re-examines important
genera across adenoma and carcinoma across study, and examines
differences and similarities between stool and tissue microbiota in the
context of colorectal tumors. Importantly, this analysis and approach
could provide valuable insights into the common genera that are both
protective and detrimental in CRC and whether broad bacterial community
measurements can account for these changes that were not provided by
earlier meta-analysis studies {[}11{]}.

Using both feces (n = 1737) and colon tissues (492 samples from 350
individuals) totaling over 2229 total samples across 14 studies
{[}12--18,21,23--27,30{]} {[}Table 1 \& 2{]}, we expand both the breadth
and scope of the previous meta-analysis to investigate whether the
bacterial community or specific members are more important risk factors
for both adenoma and carcinoma stages of disease. To accomplish this we
first assessed whether bacterial diversity changes throughout disease
(control to adenoma to carcinoma) and if it results in an increased
relative risk (RR) for adenoma or carcinoma stages of disease We then
assessed what genera, if any, increase or decrease the RR of adenoma or
carcinoma stages of disease Next, using Random Forest models, we
analyzed whether the full community or only the combined top 5 increased
and top 5 decreased RR genera resulted in better model classification,
based on the area under the curve (AUC). Finally, we also examined at
what effect and sample size the studies used were powered for and the
sample size needed to get to the traditionally accepted 80\% power. Our
results from these analyses suggests that the bacterial community
changes as disease severity worsens, that specific members are important
for disease classification, and that many of the studies are
underpowered for assessing small effect sizes.

@@

\textbf{\emph{Lower Bacterial Diversity is Associated with Increased RR
of Tumors:}} To assess differences in broad scale community metrics as
disease severity worsens Operational Taxonomic Unit (OTU) richness,
evenness, and Shannon diversity measurements were power transformed and
Z-score normalized. These metrics are commonly used to assess the total
number of OTUs, the equality of their abundance, and the overall
diversity, respectively. Using linear mixed-effect models to control for
study and variable region we assessed whether OTU richness, evenness, or
Shannon diversity changed in a step-wise manner with disease severity.
In stool, there was a significant decrease in both evenness and Shannon
diversity as disease severity moved from control to adenoma to carcinoma
(P-value = 0.025 and 0.043, respectively) {[}Figure 1A{]}. We next
tested whether the detectable differences in community significantly
increased in RR of having an adenoma or carcinoma. For fecal samples, a
decrease versus the overall median in evenness resulted in a
significantly increased RR for carcinoma (RR = 1.36 (1.15 - 1.61),
P-value = 0.00034) and adenoma (RR = 1.16 (1.01 - 1.34), P-value =
0.032) while a decrease versus the overall median in Shannon diversity
only increased the RR for carcinoma (RR = 1.33 (1.09 - 1.62), P-value =
0.0047) {[}Figure 2{]}. Using the Bray-Curtis distance metric and
PERMANOVA, it was also possible to identify significant bacterial
community changes, in specific studies, for both carcinoma-associated
and adenoma-associated microbiota versus control {[}Table S1 \& S2{]}.

@@

\textbf{\emph{Mouth-Associated Genera are Associated with an Increased
RR of Tumor:}} Next, we asked if being higher than the median relative
abundance, for any specific genera, resulted in an altered RR for
adenoma or carcinoma, in stool and colon tissue, due to our previous
observations of small increases in RR using OTU richness and Shannon
diversity. To investigate this we analyzed all common genera across each
study, in colon tissue or stool, and assessed whether a relative
abundance higher than the median results in an increase or decrease in
RR. Mouth-associated genera were commonly found in the top 5 genera
associated with an increased RR of having an adenoma
(\emph{Pyramidobacter} {[}Figure 3A{]} and \emph{Rothia} {[}Figure
3C{]}) and carcinoma (\emph{Fusobacterium}, \emph{Parvimonas},
\emph{Porphyromonas}, and \emph{Peptostreptococcus} {[}Figure 3B{]} and
\emph{Fusobacterium} {[}Figure 3D{]}) for both stool and colon tissue
samples. Conversely, genera commonly associated with a normal
gastrointestinal tract were correlated with a decreased RR for both
adenoma and carcinoma for both stool and colon tissue samples {[}Figure
3{]}. Even though mouth-associated genera were identified across disease
stage, there was little direct overlap of the top 5 increased or
decreased RR genera between both stages and sample site.

When observing RRs with a P-value less than 0.05 there was almost no
overlap between genera from stool or colon tissue and when they were
similar the RR was in opposite directions (e.g. \emph{Lactococcus})
{[}Table S8 \& S9{]}. Many of the genera that had RRs with a P-value
under 0.05 for colon tissue are also highly prevalent in contamination,
specifically, \emph{Novosphingobium}, \emph{Selemonas}, and
\emph{Achromobacter} {[}Figure 3 \& Table S8-S9{]}. For carcinoma stage
of disease, certain mouth-associated genera (\emph{Fusobacterium},
\emph{Parvimonas}, \emph{Peptostreptococcus}) had a high RR for both
colon tissue and stool samples {[}Table S10 \& S11{]}. Finally, these
data suggest that the most significantly increased RR genera for tissue
was \emph{Camplyobacter} while in stool it was \emph{Peptostreptococcus}
{[}Table S10 \& S11{]}.

\textbf{\emph{Select Community Models can Recapitulate Whole Community
Models:}} Since specific genera increased RR for carcinoma over
diversity metrics we assessed whether the bacterial community was better
at classifying disease versus only a select group of genera. We selected
these genera based on their RR and P-value significance and used two
approaches to test this question. The first approach used genus level
data and tested for differences in AUC between all genera and selected
genera. A single study was used for training the model prior to testing
on all other studies and this was repeated for every study in the
meta-analysis. The second approach used OTU level data and tested for a
generalized decrease in the 10-fold cross validation (CV) model AUC
which is a common approach used to guard against over-fitting. This was
applied across study and the AUC of the all OTUs model was compared
against the model that used only OTUs that taxonomically classified to
selected genera.

@@

Our study identifies clear differences in diversity, both at the
community level and for individual genera, present in patients with and
without CRC {[}Figure 1-3{]}. Although there was a step-wise decrease in
diversity as disease progressed from control to adenoma to carcinoma,
this did not translate into large effect sizes for the RR of tumors.
Even though mouth-associated genera increased the RR of having a
carcinoma, they did not consistently increase the RR of having an
adenoma. Additionally, our observations suggest that by combining
mouth-associated and CRC protective microbes we can classify either
adenoma or carcinoma stage of disease as well as models that use the
full community.

@@

The colon tissue-based studies did not provide a clearer understanding
of how the microbiota may be associated with tumors. Generally, the full
OTU-based models of unmatched and matched colon tissue samples were
concordant with stool samples showing that GI resident microbes were the
most prevalent in the top 10 most important variables across study
{[}Figure S7E \& F{]}. Unlike in stool, \emph{Fusobacterium} was the
only mouth-associated bacteria consistently present in the top 10 most
important variables of the full carcinoma stage models {[}Figure S7B-C
\& E-F{]}. The majority of the colon tissue-based results seem to be
study specific with many of the top 10 taxa being present only in a
single study. Additionally, the presence of genera associated with
contamination, within the top 10 most important variables for the genera
and OTU models is worrying. The low bacterial biomass of tissue samples
coupled with potential contamination could explain why these results
seem to be more sporadic than the stool results.

@@

\textbf{\emph{Matched versus Unmatched Tissue Samples:}} In general,
tissue samples with control and tumor samples from different individuals
were classified as unmatched while samples that belonged to the same
individual were classified as matched. Studies with matched data
included Burns, Dejea, Geng, and Lu while those with unmatched data were
from Burns, Flemer, Chen, and Sanapareddy. For some studies samples
became unmatched when a corresponding matched sample did not make it
through sequence processing. All samples, from both tissue sample types,
were analyzed together for the linear mixed-effect models with samples
from the same individual corrected for. For all other analysis, not
mentioned herein, matched and unmatched samples were analyzed separately
using the statistical approaches mentioned in the Statistical Analysis
section.
